\section{Discussion}
\label{sec:discuss}
In this section, we discuss two observations in preventing reward hacking and LLM's degenerate behaviors in complex end-to-end GPU programming. 

\subsection{Format check fails to prevent reward hacking}
In our early experiment, we try to use format check to prevent reward hacking in RL, which is also applied by Kevin-32B~\cite{kevin}. However, format check fails to prevent reward hacking, both in our experiments and Kevin-32B's result in Table~\ref{tab:hack}. We attribute the failure to two reasons: First, some reward hacking behaviors are very subtle and hard to be detected by a regex-based scripts, such as hard coding. Second, format check faces a trade-off, where strict checks could classify correct solution as “hacking”, and loose checks could omit some hacking solutions. Therfore, format check alone is not a good remedy for avoiding reward hacking on KernelBench. 



\subsection{Degenerate behaviors on KernelBench}
We also observe that current LLMs clearly exhibit degenerate behaviors in complex end-to-end GPU programming tasks. For example, in many Level 2/3 tasks with multiple kernels, CUDAForge and Kevin-32B show a strong tendency to only modify “easy parts,” with only minor impact on performance. For example, only write a ReLU kernel and leave all others (e.g., conv2d and GEMM) unchanged.  And they will even replace only one of multiple ReLU kernels in the reference code with a customized kernel. In this case, the program can be correct with a very minor different performance (typically less than 5\% in our experiments). A case of this kind of code generated by Kevin is shown in Appendix~\ref{app:fear}. We attribute the degenerate behaviors to two reasons:

\textbf{RL reward design pushes the model.} According to the rule-based reward design of Kevin-32B in Appendix~\ref{kevinreward}. A correct but low-speed kernel will get higher rewards than a kernel with aggressive optimizition but has a minor error, hindering model try those advanced optimization techniques. See Appendix~\ref{app:rubric} for how StitchCUDA address it.

\textbf{Current LLMs have no confidence.} We manually track the chain of thought of these models in thinking mode to figure out the reason of degenerate behaviors. We observe a lot of similar content in their chain of thought, such as “current conv2d in PyTorch is backend with cuBLAS lib, which is hard to surpass, focus on ReLU and Linear kernels first.”. This reveals that current LLMs have no confidence to challenge those real critical parts in GPU programming. And as a result, they also lose the opportunity to utilize low-level deep optimization techniques in practice, such as fine-grained tiling and tensor cores. Based on our observations, Kevin32B never use these advanced techniques in experiments, which partially explains why its speedup decreases significantly on KernelBench level 3. 





\section{Conclusion}
In this paper, we propose StitchCUDA, a multi-agent framework that integrates Rubric-based reinforcement learning with the Coder agent. We identify three main challenges in LLM-based automated end-to-end GPU program generation. We improve the multi-agent framework through well-designed multi-agent decomposition and fundamentally enhance the Coder's CUDA coding ability through rubric-based agentic reinforcement learning. Comprehensive experiment results show that StitchCUDA delivers excellent performance on end-to-end GPU programming tasks, significantly outperforming simple multi-agent approaches and RL models.